---
title: "S1 Chlorophyll-a bloom detection"
unnumbered: true
---

```{r}
#| label: setup-bloom-onset
#| include: false

# For data manipulation and plotting (loads ggplot2, dplyr, tidyr, etc.)
library(tidyverse)

# For handling and plotting spatial data
library(sf)
library(rnaturalearth)
library(ggspatial)

# For combining plots
library(cowplot)   # Used for the map insets
library(patchwork) # Used for the final density plots

# For file path management
library(here)

# For time-series calculations (rolling medians)
library(zoo)

# For creating interactive HTML tables
library(DT)


```

# S1.1 Map of Isla Isabel

A map of Isla Isabel, Mexico, is generated to provide geographical context for the study area. The figure highlights the island’s location within the eastern tropical Pacific, outlines the surrounding coastal region, and marks the island itself with a reference point.

```{r}
#| label: map_final_refinements
#| fig-width: 10
#| fig-height: 8
#| code-fold: true



# --- Define Data ---
# Define the polygon coordinates
latitudes <- c(22.1203, 22.1203, 21.5797, 21.5797, 22.1203)
longitudes <- c(-106.1763, -105.5937, -105.5937, -106.1763, -106.1763)
coords_df <- data.frame(lon = longitudes, lat = latitudes)
polygon_sf <- st_polygon(list(as.matrix(coords_df)))
polygon_sf <- st_sfc(polygon_sf, crs = 4326)

# Define the marker coordinates for Isla Isabel
marker_lat_deg <- 21 + 50/60 + 59.0/3600
marker_lon_deg <- -(105 + 52/60 + 54.0/3600)
marker_point_df <- data.frame(lon = marker_lon_deg, lat = marker_lat_deg)
marker_point_sf <- st_point(c(marker_point_df$lon, marker_point_df$lat))
marker_point_sf <- st_sfc(marker_point_sf, crs = 4326)

# Get world map data
world <- ne_countries(scale = "medium", returnclass = "sf")
coastlines <- ne_coastline(scale = "medium", returnclass = "sf")

# Create points for country labels
world_points <- cbind(world, st_coordinates(st_centroid(world$geometry)))

# --- Main Map (Zoomed Out - Pacific Coast) ---
main_map <- ggplot() +
  # TYPO FIX: Changed "antiquhite" to "antiquewhite"
  geom_sf(data = world, fill = "antiquewhite", color = "darkgray") +
  geom_sf(data = coastlines, color = "grey", size = 0.5) +
  geom_sf(data = st_as_sfc(st_bbox(polygon_sf)), fill = NA, color = "red", size = 1) +
  geom_text(data = world_points, aes(x = X, y = Y, label = name),
            color = "black", size = 3.5, fontface = "bold", check_overlap = TRUE) +
  annotate(geom = "text", x = -90, y = 25, label = "Gulf of Mexico",
           fontface = "italic", color = "steelblue4", size = 5) +
  coord_sf(xlim = c(-125, -85), ylim = c(12, 35), expand = FALSE) +
  labs(title = "", x = "Longitude (°W)", y = "Latitude (°N)") +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(face = "bold"),
    panel.grid.major = element_line(color = "lightgray", linetype = "dashed"),
    panel.background = element_rect(fill = "aliceblue")
  )

# --- Inset Map (Zoomed In on Isla Isabel) ---
inset_map <- ggplot() +
  geom_sf(data = world, fill = "antiquewhite", color = "darkgray") +
  geom_sf(data = coastlines, color = "grey", size = 0.5) +
  geom_sf(data = polygon_sf, fill = "blue", alpha = 0.5, color = "darkblue", size = 1) +
  geom_sf(data = marker_point_sf, color = "red", size = 4, shape = 17) +
  geom_text(data = marker_point_df, aes(x = lon, y = lat), label = "Isla Isabel", hjust = 0.5, vjust = -5.5, color = "black", fontface = "bold", size = 6) +
  annotation_scale(location = "bl", width_hint = 0.2) +
  annotation_north_arrow(
    location = "bl", which_north = "true",
    pad_x = unit(0.09, "in"), pad_y = unit(0.3, "in"),
    style = north_arrow_fancy_orienteering
  ) +
  coord_sf(
    xlim = c(marker_lon_deg - 0.5, marker_lon_deg + 0.5),
    ylim = c(marker_lat_deg - 0.5, marker_lat_deg + 0.5),
    expand = FALSE
  ) +
  theme_bw() +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "aliceblue"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)
  )

# --- Combine Maps using cowplot ---
mapIS<-ggdraw() +
  draw_plot(main_map) +
  draw_plot(inset_map, x = 0.05, y = 0.15, width = 0.38, height = 0.38)
mapIS
```

```{r}
ggsave(filename = here("..","Plots", "Map_Isla_Isabel.pdf"),
       plot = mapIS,
       dpi = 300, device = cairo_pdf)

ggsave(filename = here("..","Plots", "Map_Isla_Isabel.jpg"),
       plot = mapIS,
       width = 7,  
       height = 7,  
       dpi = 300)
```

# S1.2 Bloom onset

This section details the methodology used to estimate the onset of phytoplankton blooms based on satellite-derived chlorophyll-a time series.

## Chlorophyll-a bloom detection

Satellite-derived chlorophyll-a data were processed to identify the onset of seasonal blooms.

::::: panel-tabset
## Load chlorophyll data

The chlorophyll-a data is loaded from a CSV file. The data is cleaned by removing missing values (`NA`), selecting the relevant `date` and `chlorophyll` columns, and converting the date column to a proper date format using `lubridate::dmy()`

```{r}
#| label: chlorophyll data
CHL_df <- read.csv(here("..","data", "MODIS.CHL.R2022.csv"), header = TRUE) %>%
  drop_na() %>%
  dplyr::select(date, chlorophyll) %>%
  rename(daily_value = chlorophyll)

CHL_df <- CHL_df %>%
  mutate(date = dmy(date)) 
```

## Calculate sustained bloom start

Two custom functions were developed to estimate bloom onset dates for different seasonal windows. Both apply the following procedure:

-   Seasonal window: Define a “season year” spanning either December–June or September–June.

-   Baseline: Calculate the seasonal median chlorophyll-a concentration.

-   Thresholds: Set dynamic thresholds at 5%, 15%, 25%, and 30% above the seasonal median.

-   Sustained increase: Identify the first day when two consecutive 7-day rolling medians exceed the threshold, ensuring that the increase is persistent rather than transient.

-   Output: Return the bloom start date, chlorophyll-a concentration on that date, and the corresponding threshold.

:::: panel-tabset
## Winter to Spring window

This function defines the season as running from December 21st of the previous year to June 20th of the current year.

```{r}
#| label: Function to calculate sustained bloom start date

calculate_sustained_bloom_start <- function(data_frame, value_column_name = "daily_value", min_duration_days = 7) {

  thresholds <- c(0.05, 0.15,0.25, 0.30)

  # Prepare and annotate data
  data_frame <- data_frame %>%
    mutate(date = as.Date(date)) %>%
    arrange(date) %>%
    mutate(season_year = ifelse(month(date) == 12, year(date) + 1, year(date)))

  # Define seasonal window
  season_ranges <- data_frame %>%
    distinct(season_year) %>%
    mutate(
      season_start_date = ymd(paste0(season_year - 1, "-12-21")),
      season_end_date = ymd(paste0(season_year, "-06-20"))
    )

  # Filter seasonal data
  seasonal_data <- data_frame %>%
    left_join(season_ranges, by = "season_year") %>%
    filter(date >= season_start_date & date <= season_end_date) %>%
    arrange(season_year, date)

  # Compute yearly medians
  medians <- seasonal_data %>%
    group_by(season_year) %>%
    summarise(yearly_median = median(!!sym(value_column_name), na.rm = TRUE), .groups = "drop")

  # Initialize final result with medians
  all_results <- medians
  bloom_date_raws <- list()

  for (thr in thresholds) {
    label <- paste0("thr", thr * 100)

    # Add threshold column
    threshold_tbl <- medians %>%
      mutate(threshold = yearly_median * (1 + thr))

    seasonal_thr_data <- seasonal_data %>%
      left_join(threshold_tbl, by = "season_year")

    # Compute rolling medians
    seasonal_thr_data <- seasonal_thr_data %>%
      group_by(season_year) %>%
      mutate(
        rolling_median_week1 = rollapply(
          !!sym(value_column_name), min_duration_days, median, na.rm = TRUE, fill = NA, align = "left"
        ),
        rolling_median_week2 = rollapply(
          lead(!!sym(value_column_name), n = min_duration_days), min_duration_days, median, na.rm = TRUE, fill = NA, align = "left"
        ),
        cond_i = !is.na(rolling_median_week1) & rolling_median_week1 >= threshold,
        cond_ii = !is.na(rolling_median_week2) & rolling_median_week2 >= threshold
      ) %>%
      filter(cond_i & cond_ii) %>%
      summarise(bloom_date = min(date, na.rm = TRUE), .groups = "drop")

    # Store raw date for difference calculation
    bloom_date_raws[[label]] <- seasonal_thr_data %>%
      rename(!!paste0("bloom_date_raw_", label) := bloom_date)

    # Join back to get bloom values and threshold
    enriched <- seasonal_thr_data %>%
      left_join(seasonal_data, by = "season_year") %>%
      group_by(season_year) %>%
      summarise(
        !!paste0("bloom_start_", label) := format(first(bloom_date), "%d/%m/%Y"),
        !!paste0("value_at_bloom_", label) := seasonal_data[[value_column_name]][match(first(bloom_date), seasonal_data$date)],
        !!paste0("threshold_", label) := medians$yearly_median[medians$season_year == first(season_year)] * (1 + thr),
        .groups = "drop"
      )

    # Join enriched info to results
    all_results <- all_results %>%
      left_join(enriched, by = "season_year")
  }

  # Join all bloom date raws
  bloom_date_df <- reduce(bloom_date_raws, left_join, by = "season_year")

  # Add differences between 5% vs 15% and 30%
  all_results <- all_results %>%
    left_join(bloom_date_df, by = "season_year") %>%
    mutate(
      diff_days_5_vs_15 = as.numeric(bloom_date_raw_thr15 - bloom_date_raw_thr5),
      diff_days_5_vs_30 = as.numeric(bloom_date_raw_thr30 - bloom_date_raw_thr5)
    ) %>%
    select(-starts_with("bloom_date_raw_"))

  return(all_results)
}
 bloom_startsA <- calculate_sustained_bloom_start(CHL_df)
#write.csv(bloom_starts2, here("data", "bloom_dates.csv"), row.names = FALSE)
```

## Autumn to Spring Window

This function defines a broader season, running from September 21st of the previous year to June 20th of the current year.

```{r}
#| label: Function to calculate sustained bloom start from Autumn to Spring


calculate_sustained_bloom_startB <- function(data_frame, value_column_name = "daily_value", min_duration_days = 7) {

  thresholds <- c(0.05, 0.15,0.25, 0.30)

  # Prepare and annotate data
  data_frame <- data_frame %>%
    mutate(date = as.Date(date)) %>%
    arrange(date) %>%
    mutate(season_year = ifelse(month(date) == 9, year(date) + 1, year(date)))

  # Define seasonal window
  season_ranges <- data_frame %>%
    distinct(season_year) %>%
    mutate(
      season_start_date = ymd(paste0(season_year - 1, "-09-21")),
      season_end_date = ymd(paste0(season_year, "-06-20"))
    )

  # Filter seasonal data
  seasonal_data <- data_frame %>%
    left_join(season_ranges, by = "season_year") %>%
    filter(date >= season_start_date & date <= season_end_date) %>%
    arrange(season_year, date)

  # Compute yearly medians
  medians <- seasonal_data %>%
    group_by(season_year) %>%
    summarise(yearly_median = median(!!sym(value_column_name), na.rm = TRUE), .groups = "drop")

  # Initialize final result with medians
  all_results <- medians
  bloom_date_raws <- list()

  for (thr in thresholds) {
    label <- paste0("thr", thr * 100)

    # Add threshold column
    threshold_tbl <- medians %>%
      mutate(threshold = yearly_median * (1 + thr))

    seasonal_thr_data <- seasonal_data %>%
      left_join(threshold_tbl, by = "season_year")

    # Compute rolling medians
    seasonal_thr_data <- seasonal_thr_data %>%
      group_by(season_year) %>%
      mutate(
        rolling_median_week1 = rollapply(
          !!sym(value_column_name), min_duration_days, median, na.rm = TRUE, fill = NA, align = "left"
        ),
        rolling_median_week2 = rollapply(
          lead(!!sym(value_column_name), n = min_duration_days), min_duration_days, median, na.rm = TRUE, fill = NA, align = "left"
        ),
        cond_i = !is.na(rolling_median_week1) & rolling_median_week1 >= threshold,
        cond_ii = !is.na(rolling_median_week2) & rolling_median_week2 >= threshold
      ) %>%
      filter(cond_i & cond_ii) %>%
      summarise(bloom_date = min(date, na.rm = TRUE), .groups = "drop")

    # Store raw date for difference calculation
    bloom_date_raws[[label]] <- seasonal_thr_data %>%
      rename(!!paste0("bloom_date_raw_", label) := bloom_date)

    # Join back to get bloom values and threshold
    enriched <- seasonal_thr_data %>%
      left_join(seasonal_data, by = "season_year") %>%
      group_by(season_year) %>%
      summarise(
        !!paste0("bloom_start_", label) := format(first(bloom_date), "%d/%m/%Y"),
        !!paste0("value_at_bloom_", label) := seasonal_data[[value_column_name]][match(first(bloom_date), seasonal_data$date)],
        !!paste0("threshold_", label) := medians$yearly_median[medians$season_year == first(season_year)] * (1 + thr),
        .groups = "drop"
      )

    # Join enriched info to results
    all_results <- all_results %>%
      left_join(enriched, by = "season_year")
  }

  # Join all bloom date raws
  bloom_date_df <- reduce(bloom_date_raws, left_join, by = "season_year")

  # Add differences between 5% vs 15% and 30%
  all_results <- all_results %>%
    left_join(bloom_date_df, by = "season_year") %>%
    mutate(
      diff_days_5_vs_15 = as.numeric(bloom_date_raw_thr15 - bloom_date_raw_thr5),
      diff_days_5_vs_30 = as.numeric(bloom_date_raw_thr30 - bloom_date_raw_thr5)
    ) %>%
    select(-starts_with("bloom_date_raw_"))

  return(all_results)
}
bloom_startsB <- calculate_sustained_bloom_startB(CHL_df)
#write.csv(bloom_startsB, here("data", "bloom_datesB.csv"), row.names = FALSE)
```

## Summary of bloom start dates

::: panel-tabset
## Summary for Winter to Spring

This table summarizes bloom start dates detected within the December–June seasonal window.

```{r}
#| label: Summary A
#| echo: false
DT::datatable(
  bloom_startsA,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    dom = 'Bfrtip',
    buttons = c('copy', 'csv')
  ),
  extensions = 'Buttons',
  rownames = FALSE
) %>%
  formatRound(columns = c("yearly_median","value_at_bloom_thr5", "threshold_thr5","value_at_bloom_thr15","threshold_thr15","value_at_bloom_thr25","threshold_thr25","value_at_bloom_thr30","threshold_thr30"), digits = 3)%>%
  formatStyle(columns = names(bloom_startsA), `text-align` = 'center')
```

## Summary for Autumn to Spring

This table summarizes bloom start dates detected within the September–June seasonal window.

```{r}
#| label: Summary B
#| echo: false
DT::datatable(
  bloom_startsB,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    dom = 'Bfrtip',
    buttons = c('copy', 'csv')
  ),
  extensions = 'Buttons',
  rownames = FALSE
) %>%
  formatRound(columns = c("yearly_median","value_at_bloom_thr5", "threshold_thr5","value_at_bloom_thr15","threshold_thr15","value_at_bloom_thr25","threshold_thr25","value_at_bloom_thr30","threshold_thr30"), digits = 3)%>%
  formatStyle(columns = names(bloom_startsB), `text-align` = 'center')
```
:::
::::
:::::

# S1.3 Visualization of Chlorophyl-a bloom start dates

To examine the distribution and timing of bloom onsets, detection results were transformed and visualized. Density plots compare bloom start dates across thresholds and seasonal windows.

### Data processing for visualization

Prior to plotting, results were transformed into long format for analysis with ggplot2. A helper function (process_for_plotting) standardized bloom dates across years and summarized them by detection threshold. Dates from September–December were aligned to 2000 and January–June dates to 2001, creating a continuous seasonal year.

```{r}
#| label: Frequency of bloom start dates
#| code-fold: true

process_for_plotting <- function(results) {
  results %>%
    select(season_year, starts_with("bloom_start_")) %>%
    pivot_longer(
      cols = -season_year,
      names_to = "threshold_id",
      values_to = "bloom_date_str",
      values_drop_na = TRUE
    ) %>%
    mutate(
      threshold_label = gsub("bloom_start_thr", "", threshold_id),
      threshold = factor(paste0(threshold_label, "%"), levels = c("5%", "15%", "25%", "30%")),
      bloom_date = dmy(bloom_date_str),
      bloom_day_month = if_else(
          month(bloom_date) >= 9,
          update(bloom_date, year = 2000),
          update(bloom_date, year = 2001)
      )
    )
}

bloom_dates_long_A <- process_for_plotting(bloom_startsA)
bloom_dates_long_B <- process_for_plotting(bloom_startsB)


sample_sizes_A <- bloom_dates_long_A %>%
  group_by(threshold) %>%
  summarise(label = paste("n =", n_distinct(season_year)), .groups = "drop")

sample_sizes_B <- bloom_dates_long_B %>%
  group_by(threshold) %>%
  summarise(label = paste("n =", n_distinct(season_year)), .groups = "drop")


# Calculate mean dates for vertical lines
mean_dates_A <- bloom_dates_long_A %>%
  group_by(threshold) %>%
  summarise(mean_date = as.Date(mean(as.numeric(bloom_day_month)), origin = "1970-01-01"), .groups = "drop")

mean_dates_B <- bloom_dates_long_B %>%
  group_by(threshold) %>%
  summarise(mean_date = as.Date(mean(as.numeric(bloom_day_month)), origin = "1970-01-01"), .groups = "drop")



```

### Generating and combining density plots

Density plots were generated for each seasonal window, faceted by threshold level. Dashed vertical lines mark the mean bloom onset, and sample sizes are displayed in each panel. The two plots were combined for side-by-side comparison.

```{r}
#| label: Area plot of bloom start dates

# --- Define the color palette once ---
bloom_colors <- c("5%"="#2c7bb6", "15%"="#abd9e9", "25%"="#fdae61", "30%"="#d7191c")

# --- Plot A ---
area_plot_A <- ggplot(bloom_dates_long_A, 
                      # Add 'color' aesthetic
                      aes(x = bloom_day_month, fill = threshold, color = threshold)) +
  
  # --- Replace geom_histogram with geom_area ---
  geom_area(
    stat = "bin",
    binwidth = 1,         # Keeps the 7-day binning
    boundary = 0,
    alpha = 0.6,          # Adjusted alpha for better overlap
    position = "identity",
    linewidth = 0.5       # Controls outline thickness
  ) +
  
  geom_text(data = sample_sizes_A, aes(x = as.Date("2001-06-15"), y = Inf, label = label), vjust = 1.5, hjust = 1, size = 3, inherit.aes = FALSE) +
  facet_wrap(~threshold, ncol = 1, scales = "fixed") +
  labs(subtitle = "Winter to Spring", x = NULL, y = "Count") +
  theme_light() +
  scale_fill_manual(values = bloom_colors) +
  scale_color_manual(values = bloom_colors) + # <-- Add matching color scale
  scale_x_date(limits = c(as.Date("2000-09-01"), as.Date("2001-06-20")), date_labels = "%b", date_breaks = "1 month") +
  theme(plot.subtitle = element_text(hjust = 0.5), legend.position = "none", strip.text = element_text(face = "bold"))

# --- Plot B ---
area_plot_B <- ggplot(bloom_dates_long_B, 
                      # Add 'color' aesthetic
                      aes(x = bloom_day_month, fill = threshold, color = threshold)) +
  
  # --- Replace geom_histogram with geom_area ---
  geom_area(
    stat = "bin",
    binwidth = 1,         # Keeps the 7-day binning
    boundary = 0,
    alpha = 0.6,
    position = "identity",
    linewidth = 0.5
  ) +
  
  facet_wrap(~threshold, ncol = 1, scales = "fixed") + 
  geom_text(data = sample_sizes_B, aes(x = as.Date("2001-06-15"), y = Inf, label = label), vjust = 1.5, hjust = 1, size = 3, inherit.aes = FALSE)+
  labs(subtitle = "Autumn to Spring", x = NULL, y = "Count") +
  theme_light() +
  scale_fill_manual(values = bloom_colors) +
  scale_color_manual(values = bloom_colors) + # <-- Add matching color scale
  scale_x_date(limits = c(as.Date("2000-09-01"), as.Date("2001-06-20")), date_labels = "%b", date_breaks = "1 month") +
  theme(plot.subtitle = element_text(hjust = 0.5), legend.position = "none", strip.text = element_text(face = "bold"))

# --- Combine plots ---
combined_plot <- area_plot_A + area_plot_B +
  plot_annotation(
    tag_levels = 'A'
  )

# Print the final combined plot
print(combined_plot)
```

```{r}
#| label: Histograms of bloom start dates
#| code-fold: true


histogram_plot_A <- ggplot(bloom_dates_long_A, aes(x = bloom_day_month, fill = threshold)) +
  geom_histogram(
    binwidth = 7,         # Sets bin width to 7 days (1 week)
    boundary = 0,         # Aligns bins cleanly with the week
    alpha = 0.7, 
    position = "identity" # Allows transparent overlap
  ) +
  geom_text(data = sample_sizes_A, aes(x = as.Date("2001-06-15"), y = Inf, label = label), vjust = 1.5, hjust = 1, size = 3, inherit.aes = FALSE) +
  facet_wrap(~threshold, ncol = 1, scales = "fixed") +
  labs(subtitle = "Winter to Spring", x = NULL, y = "Count") +
  theme_light() +
  scale_fill_manual(values = c("5%"="#2c7bb6", "15%"="#abd9e9", "25%"="#fdae61", "30%"="#d7191c")) +
  scale_x_date(limits = c(as.Date("2000-09-01"), as.Date("2001-06-20")), date_labels = "%b", date_breaks = "1 month") +
  theme(plot.subtitle = element_text(hjust = 0.5), legend.position = "none",strip.text = element_text(face = "bold"))


histogram_plot_B <- ggplot(bloom_dates_long_B, aes(x = bloom_day_month, fill = threshold)) +
  geom_histogram(
    binwidth = 7,         # Sets bin width to 7 days (1 week)
    boundary = 0,         # Aligns bins cleanly with the week
    alpha = 0.7, 
    position = "identity" # Allows transparent overlap
  ) +
  facet_wrap(~threshold, ncol = 1, scales = "fixed") + 
  geom_text(data = sample_sizes_B, aes(x = as.Date("2001-06-15"), y = Inf, label = label), vjust = 1.5, hjust = 1, size = 3, inherit.aes = FALSE)+
  labs(subtitle = "Autumn to Spring", x = NULL, y = "Count") +
  theme_light() +
  scale_fill_manual(values = c("5%"="#2c7bb6", "15%"="#abd9e9", "25%"="#fdae61", "30%"="#d7191c")) +
  scale_x_date(limits = c(as.Date("2000-09-01"), as.Date("2001-06-20")), date_labels = "%b", date_breaks = "1 month") +
  theme(plot.subtitle = element_text(hjust = 0.5), legend.position = "none",strip.text = element_text(face = "bold"))

# Combine plots with patchwork
combined_plot <- histogram_plot_A  +histogram_plot_B  +
  plot_annotation(
    tag_levels = 'A'
  )

# Print the final combined plot
print(combined_plot)

```

```{r}
ggsave(filename = here("..","Plots", "bloom_thresholds.pdf"),
       plot = combined_plot,
       dpi = 300, device = cairo_pdf)

ggsave(filename = here("..","Plots", "bloom_thresholds.jpg"),
       plot = combined_plot,
       width = 7,  
       height = 7,  
       dpi = 300)
```

### Interpretation of Bloom start date figure

This figure shows the frequency distribution of phytoplankton bloom start dates. The left column displays results for a Winter–Spring seasonal window (Dec–Jun), while the right column uses a broader Autumn–Spring window (Sep–Jun). Each row represents a different detection threshold (5%, 15%, 25%, and 30%) relative to the seasonal median chlorophyll-a concentration.

Key patterns include:

Higher Thresholds Delay Bloom Onset: As the detection threshold increases from 5% to 30%, the timing of the bloom onset is consistently pushed later into the year. The earliest blooms, detected with the 5% threshold, peak in December and January. In contrast, the most intense blooms (30% threshold) do not typically start until late January or February. This pattern is expected, as it takes more time for phytoplankton biomass to accumulate to higher concentrations.

Bimodal Onset: The Autumn-Spring analysis reveals a distinct bimodal pattern in bloom timing. An initial wave of blooms occurs in early autumn (September-October), followed by a lull. After this quiet period, the primary, more sustained spring bloom begins in winter, with its subsequent timing and frequency closely mirroring the patterns observed in the Winter-Spring-only analysis. This overall bimodal structure strongly suggests the early autumn blooms are event-driven phenomena, likely triggered by nutrient injection from late-season storms or hurricanes.

Bloom variability: The decreasing sample size (n) for higher thresholds indicates strong year-to-year variability. This suggests that high-magnitude blooms are not an annual occurrence, pointing to the influence of variable environmental conditions.

::: {.callout-tip appearance="simple" icon="false"}
```{r}
#| echo: false
#| warning: false
#| message: false
utils::sessionInfo()
```
:::
