---
title: "S2. Phenological mismatches"
---

```{r}
#| label: setup-pheno-mismatch
#| include: false

# For data manipulation and plotting (loads ggplot2, dplyr, tidyr, etc.)
library(tidyverse)

# For file path management
library(here)
```

This section examines phenological mismatches, defined as the temporal difference between breeding events and the onset of phytoplankton blooms.

## Data preparation

Data on phenological mismatches were compiled and prepared for visualization. The dataset includes mismatches calculated separately for males and females, across bloom detection thresholds of 5%, 15%, and 25%, and for two seasonal windows (Winter–Spring and Autumn–Spring).

The data were restructured for plotting in `ggplot2`, producing a combined dataset used to generate density plots. These plots are faceted by sex and season, and color-coded by threshold level.

## Mismatch visualization

Phenological mismatches were visualized as density distributions. Each panel corresponds to a combination of sex (females or males) and seasonal window (Winter–Spring or Autumn–Spring), with distributions compared across thresholds (5%, 15%, and 25%).

```{r}
#| label: Plot pheno mismatch
#| code-fold: true
dat <- read.csv(here::here("..","data", "Lay_date_mismatches_250702.csv"))


# 1. Prepare 'Winter to Spring' data
data_A <- dat %>%
  select(SEX, X5MISMATCH, X15MISMATCH, X25MISMATCH) %>%
  rename(
    `5% threshold` = `X5MISMATCH`,
    `15% threshold` = `X15MISMATCH`,
    `25% threshold` = `X25MISMATCH`
  ) %>%
  pivot_longer(
    cols = -SEX,
    names_to = "Mismatch_Type",
    values_to = "Value"
  ) %>%
  mutate(Season = "Winter to Spring")

# 2. Prepare 'Autumn to Spring' data
data_B <- dat %>%
  select(SEX, X5MISMATCHB, X15MISMATCHB) %>%
  rename(
    `5% threshold` = `X5MISMATCHB`,
    `15% threshold` = `X15MISMATCHB`
  ) %>%
  pivot_longer(
    cols = -SEX,
    names_to = "Mismatch_Type",
    values_to = "Value"
  ) %>%
  mutate(Season = "Autumn to Spring")


combined_data <- bind_rows(data_A, data_B) %>%
  mutate(

    Sex_Label = factor(if_else(SEX == 0, "Males", "Females"), levels = c("Females", "Males")),

    Season = factor(Season, levels = c("Winter to Spring", "Autumn to Spring")),

    Mismatch_Type = factor(Mismatch_Type, levels = c("5% threshold", "15% threshold", "25% threshold"))
  ) %>%

  filter(!is.na(Value))



```

```{r}
# --- NEW 2-COLOR PALETTE ---
season_palette <- c(
  "Winter to Spring" = "#0072B2", # Blue
  "Autumn to Spring" = "#DAA520"  # Gold (Goldenrod)
)

# --- MODIFIED PLOT CODE ---

# Map fill and color to 'Season'
mismatch_plot <- ggplot(combined_data, aes(x = Value, 
                                          fill = Season, 
                                          color = Season)) +
  
  geom_area(
    stat = "bin",
    binwidth = 1,
    position = "identity",
    alpha = 0.5, # Alpha is critical for seeing the overlap
    linewidth = 0
  ) +
  
  # Faceting remains the same (3 rows, 2 cols)
 facet_grid(Mismatch_Type ~ Sex_Label) +
  
 labs(
  title = "",
  x = "Phenological mismatch (days)",
  y = "Count"
 ) +
  
  # --- Use the new 2-color palette ---
 scale_fill_manual(
    name = "Season", 
    values = season_palette, 
    drop = FALSE
  ) +
  scale_color_manual(
    name = "Season", 
    values = season_palette, 
    drop = FALSE
  ) +
  
 theme_light() +
 
 scale_x_continuous(limits = c(-150, 250), breaks = seq(-150, 250, 50)) +
 
  theme(
  legend.position = "right",
    strip.text = element_text(face = "bold")
  )

print(mismatch_plot)
```

```{r}
ggsave(filename = here("..","Plots", "mismatch_thresholds.pdf"),
       plot = mismatch_plot,
       dpi = 300, device = cairo_pdf)

ggsave(filename = here("..","Plots", "mismatch_thresholds.jpg"),
       plot = mismatch_plot,
       width = 9,  
       height = 7,  
       dpi = 300)
```

## Interpretation of mismatch figure

-   The density plots reveal a consistent bimodal distribution across sexes, thresholds, and seasonal windows.

-   Primary mode: Centered at approximately 0 days, indicating synchrony between breeding and bloom onset.

-   Secondary mode: A smaller peak around +100 days, reflecting cases where breeding occurs long after bloom initiation.

-   Sex and season effects: The distributions are highly similar between males and females and across seasonal windows, suggesting that mismatch estimates are robust to sex-specific or seasonal definitions.

::: {.callout-tip appearance="simple" icon="false"}
```{r}
#| echo: false
#| warning: false
#| message: false
utils::sessionInfo()
```
:::
