---
title: "S4. Model diagnostics and selection"
subtitle: "Diagnostic checks and leave-one-out model comparison"
unnumbered: true
toc: true
toc-depth: 3
---

::: lead
This chapter documents the workflow used to evaluate model diagnostics and compare fitted Bayesian models. It includes automated checks for common fitting problems, such as divergent transitions and low effective sample size, together with predictive model comparison using leave-one-out cross-validation.
:::

::: {.callout-note appearance="simple"}
## How to read this chapter

This page is intended as a reproducible workflow record for diagnostic assessment and model selection. Tables shown here summarize pre-computed outputs used to identify problematic models and rank supported candidates within each analysis set.
:::

```{r}
#| label: setup-model-diagnostics
#| include: false

# For data manipulation (loads dplyr, tibble, etc.)
library(tidyverse)

# For file path management
library(here)

# For accessing model diagnostics (e.g., divergences)
library(brms)

# For summarizing posterior draws (e.g., for ESS checks)
library(posterior)

# For model comparison using LOO-CV
library(loo)

# For creating interactive HTML tables
library(DT)
```

## S4.1 Overview

Two complementary procedures were applied to the pheno_mismatch and fleds_full analysis sets:

-   Model diagnostics: automated checks for divergent transitions and low effective sample size (ESS)

-   Model comparison: ranking of supported models using leave-one-out cross-validation (LOO-CV)

::: {.callout-important appearance="simple"}
Diagnostics for bloom onset location–scale models were inspected directly from model summaries because only one model was fitted per threshold.
:::

## S4.2 Diagnostic and comparison functions

The workflow relied on three helper functions: one to summarize diagnostics, one to inventory saved LOO objects, and one to compare models separately by sex.

::: panel-tabset
## `create_diagnostic_summary()`

This function automates model-diagnostic checks for saved Bayesian model objects. It searches a target analysis directory, loads model files, and performs one of two checks:

-   counts divergent transitions, which may indicate problematic posterior exploration

-   flags parameters with low bulk or tail ESS, using a conservative threshold of 400

The output is a summary table listing the sex, model name, diagnostic result, and any problematic parameters.

```{r}
#| label: create-diagnostic-summary
#| eval: false
#| code-fold: true
#| code-summary: "Show diagnostic summary function"
create_diagnostic_summary <- function(analysis_name, check_type = "divergences", ess_threshold = 400) {
  # --- Load required libraries ---
  if (!require(here)) {install.packages("here"); library(here)}
  if (!require(brms)) {install.packages("brms"); library(brms)}
  if (!require(dplyr)) {install.packages("dplyr"); library(dplyr)}

  # --- 1. Find all male and female model files ---
  output_dir <- here("Rdata", analysis_name)
  model_files <- list.files(output_dir, pattern = "^(male|female)_fit_.*\\.rds$", full.names = TRUE)

  if (length(model_files) == 0) {
    message("No model files found in ", output_dir)
    return(invisible(NULL))
  }

  # --- 2. Initialize a list to store results ---
  results_list <- list()

  # --- 3. Loop through files, run checks, and store results ---
  for (model_path in model_files) {
    fit <- readRDS(model_path)
    file_name <- basename(model_path)
    
    # Extract sex and model name
    sex <- sub("_fit_.*", "", file_name)
    model_name <- sub("^(male|female)_fit_(.*)\\.rds$", "\\2", file_name)

    message(paste("Checking:", sex, model_name))

    # --- Run the requested check ---
    if (tolower(check_type) == "divergences") {
      nuts_params <- nuts_params(fit)
      check_value <- sum(subset(nuts_params, Parameter == "divergent__")$Value)
      problem_params <- ifelse(check_value > 0, "Divergences present", NA_character_)

    } else if (tolower(check_type) == "ess") {
      if (!require(posterior)) {install.packages("posterior"); library(posterior)}
      draws_summary <- posterior::summarise_draws(fit)
      low_ess_params <- filter(draws_summary, ess_bulk < ess_threshold | ess_tail < ess_threshold)
      check_value <- nrow(low_ess_params)
      problem_params <- ifelse(check_value > 0, paste(low_ess_params$variable, collapse = ", "), NA_character_)

    } else {
      stop("Invalid 'check_type'. Please use 'divergences' or 'ess'.")
    }

    # Store the results for the current model
    results_list[[length(results_list) + 1]] <- tibble(
      Sex = sex,
      Model = model_name,
      Check_Result = check_value,
      Problematic_Parameters = problem_params
    )
  }

  # --- 4. Combine list into a final table and return ---
  final_table <- bind_rows(results_list)

  # Rename the Check_Result column to be more descriptive
  if (tolower(check_type) == "divergences") {
    final_table <- rename(final_table, Divergences = Check_Result)
  } else {
    final_table <- rename(final_table, Low_ESS_Count = Check_Result)
  }
  
  return(final_table)
}
```

## `list_analysis_models()`

This function inventories saved LOO objects for a given analysis and returns a table listing the sex, model name, and file path for each available model.

```{r}
#| label: list-analysis-models
#| eval: false
#| code-fold: true
#| code-summary: "Show model inventory function"
list_analysis_models <- function(analysis_name) {
  if (!require(here)) {install.packages("here"); library(here)}
  if (!require(dplyr)) {install.packages("dplyr"); library(dplyr)}
  
  output_dir <- here("Rdata", analysis_name)
  all_loo_files <- list.files(output_dir, pattern = "^(male|female)_loo_.*\\.rds$", full.names = TRUE)
  
  if (length(all_loo_files) == 0) {
    message("No 'male' or 'female' LOO files found in ", output_dir)
    return(tibble())
  }
  
  model_inventory <- tibble(file_path = all_loo_files) %>%
    mutate(
      file_name = basename(file_path),
      sex = sub("_loo_.*", "", file_name),
      model_name = sub("^(male|female)_loo_(.*)\\.rds$", "\\2", file_name)
    ) %>%
    select(sex, model_name, file_path)
    
  return(model_inventory)
}
```

## `compare_models_by_sex()`

This function compares models using the loo package. Models are grouped by sex, optionally excluding problematic fits, and ranked by expected log predictive density (ELPD). The resulting output includes ELPD differences, standard errors, and LOOIC-based comparisons.

```{r}
#| label: compare-models-by-sex
#| eval: false
#| code-fold: true
#| code-summary: "Show LOO comparison function"

compare_models_by_sex <- function(model_inventory, exclude_models = NULL) {
  if (!require(loo)) {install.packages("loo"); library(loo)}
  if (!require(dplyr)) {install.packages("dplyr"); library(dplyr)}

  comparison_results <- list()

  for (current_sex in c("male", "female")) {
    cat(paste0("\n\n###--- LOO Comparison for ", toupper(current_sex), " Models ---###\n\n"))

    # Filter the inventory for the current sex
    sex_models_df <- model_inventory %>% filter(sex == current_sex)

    # Get the specific list of models to exclude
    models_to_exclude <- exclude_models[[current_sex]]
    
    if (!is.null(models_to_exclude) && nrow(sex_models_df) > 0) {
      sex_models_df <- sex_models_df %>% filter(!model_name %in% models_to_exclude)
      message(paste("For", current_sex, ", excluding models:", paste(models_to_exclude, collapse=", ")))
    }

    if (nrow(sex_models_df) < 2) {
      message(paste("Not enough models to compare for '", current_sex, "' (need at least 2). Skipping."))
      next
    }

    # Load the relevant LOO objects into a named list
    loo_list <- setNames(
      lapply(sex_models_df$file_path, readRDS), 
      sex_models_df$model_name
    )

    message(paste0("Comparing ", length(loo_list), " '", current_sex, "' models..."))
    
    # Perform the comparison and print it
    comparison <- loo_compare(x = loo_list)
    print(comparison)
    
    comparison_results[[current_sex]] <- comparison
  }

  invisible(comparison_results)
}
```
:::

## S4.3 Application to pheno_mismach

This section applies the diagnostic and model-comparison workflow to the phenological mismatch analysis set.

:::: panel-tabset
## Diagnostic Checks

Both divergent transitions and low ESS values were evaluated across all fitted models in the pheno_mismatch analysis set.

::: panel-tabset
## Divergent transitions

```{r}
#| label: run-divergence-checks-pm
#| eval: false
#| code-fold: true
#| code-summary: "Show divergence diagnostic code"

# Check for divergent transitions
divergence_summaryPM <- create_diagnostic_summary("pheno_mismatch", check_type = "divergences")
print(divergence_summaryPM)
```

```{r}
#| label: DIVERGENCIES PM
#| echo: false
divergence_summaryPM<-read.csv(here("summaries","divergence_summaryPM.csv"),header=T)
DT::datatable(divergence_summaryPM, options = list(pageLength = 5))

```

## Effective Sample Size (ESS)

```{r}
#| label: run-ess-checks-pm
#| eval: false
#| code-fold: true
#| code-summary: "Show ESS diagnostic code"
# Check for low Effective Sample Size (ESS)
ess_summaryPM <- create_diagnostic_summary("pheno_mismatch", check_type = "ess", ess_threshold = 400)
print(ess_summaryPM)
```

```{r}
#| label: ESS PM
#| echo: false
ess_summaryPM<-read.csv(here("summaries","ess_summaryPM.csv"),header=T)
DT::datatable(ess_summaryPM, options = list(pageLength = 5))
```
:::
::::

# Model comparison

Pre-computed LOO objects were compared separately for females and males after excluding models flagged as problematic in the diagnostic checks.

```{r}
#| label: run-model-comparison-pm
#| eval: false
#| code-fold: true
#| code-summary: "Show LOO comparison code for pheno_mismatch"
# Create an inventory of all LOO objects for the analysis
model_inventoryPM <- list_analysis_models("pheno_mismatch")

# Define models to exclude from comparison (e.g., due to divergences or low ESS)
bad_models_list <- list(
  male = c("m12","m6","m9"),
   female = c("m12","m6","m9")
)
# Run the LOO comparison, excluding the problematic models
all_comparisonsPM <- compare_models_by_sex(model_inventoryPM, exclude_models = bad_models_list)
```

::: panel-tabset
## Female summary

```{r}
#| label: MODELS PM females
#| echo: false
all_comparisonsPMf<-read.csv(here("summaries","all_comparisonsPMf.csv"),header=T) %>%
  select(Model, elpd_diff, se_diff)
DT::datatable(all_comparisonsPMf, options = list(pageLength = 5))

```

## Male summary

```{r}
#| label: MODELS PM males
#| echo: false
all_comparisonsPMm<-read.csv(here("summaries","all_comparisonsPMm.csv"),header=T) %>%
  select(Model, elpd_diff, se_diff)
DT::datatable(all_comparisonsPMm, options = list(pageLength = 5))

```
:::

## S4.4 Number of fledglings (`fleds_full`)

::: panel-tabset
## Diagnostic Checks

Both divergent transitions and low ESS values were evaluated across all fitted models in the fleds_full analysis set.

::: panel-tabset

## Divergent transitions

```{r}
#| label: run-divergence-checks-fs
#| eval: false
#| code-fold: true
#| code-summary: "Show divergence diagnostic code for fleds_full"

divergence_summaryFS <- create_diagnostic_summary("fleds_full", check_type = "divergences")
DT::datatable(divergence_summaryFS, options = list(pageLength = 5))


```

```{r}
#| label: divergences fleds_full
#| echo: false
divergence_summaryFS <-read.csv(here("summaries","divergence_summaryFS.csv"),header=T)

DT::datatable(divergence_summaryFS, options = list(pageLength = 5))
```

## Effective sample size (ESS)

```{r}
#| label: run-ess-checks-fs
#| eval: false
#| code-fold: true
#| code-summary: "Show ESS diagnostic code for fleds_full"
ess_summaryFS <- create_diagnostic_summary("fleds_full", check_type = "ess", ess_threshold = 400)
DT::datatable(ess_summaryFS, options = list(pageLength = 5))
```

```{r}
#| label: ESS fleds_full
#| echo: false
ess_summaryFS<-read.csv(here("summaries","ess_summaryFS .csv"),header=T)
DT::datatable(ess_summaryFS, options = list(pageLength = 5))
```
:::

## Model comparison

Saved LOO objects were compared separately for females and males after excluding models flagged as problematic in the diagnostic stage.

```{r}
#| label: run-model-comparison-fs
#| eval: false
#| code-fold: true
#| code-summary: "Show LOO comparison code for fleds_full"

# Create an inventory of all LOO objects for the analysis
model_inventoryFS <- list_analysis_models("fleds_full")
print(model_inventoryFS)

# Define models to exclude from comparison
bad_models_listFS <- list(
  male = c("ABS10","ABS16","Q3"
           ),
   female = c("ABS3","m3","Q9","Q3","Q6"
              )
)

# Run the LOO comparison
all_comparisonsFS <- compare_models_by_sex(model_inventoryFS, exclude_models = bad_models_listFS)
```

::: panel-tabset
## Female summary

```{r}
#| label: MODELS FS female
#| echo: false
all_comparisonsFSf <-read.csv(here("summaries","all_comparisonsFSf.csv"),header=T) %>%
  select(Model, elpd_diff, se_diff)
DT::datatable(all_comparisonsFSf, options = list(pageLength = 5))
```

## Male summary

```{r}
#| label: MODELS FS male
#| echo: false
all_comparisonsFSm <-read.csv(here("summaries","all_comparisonsFSm.csv"),header=T) %>%
  select(Model, elpd_diff, se_diff)
DT::datatable(all_comparisonsFSm, options = list(pageLength = 5))
```
:::

## Reproducibility

::: {.callout-tip appearance="simple" icon="false"}
## Session information

```{r}
#| echo: false
#| warning: false
#| message: false
utils::sessionInfo()
```
:::
