---
title: " S5. Model comparison"
---

```{r}
#| label: setup-model-comparison
#| include: false

# Load libraries needed for this chapter
pacman::p_load(here, dplyr, DT,loo)
```

This section presents the results of the model comparisons used to identify the best-performing model for each analysis.

All candidate models were evaluated with Leave-One-Out Cross-Validation (LOO-CV), which estimates a model’s out-of-sample predictive accuracy. Models are ranked by their expected log predictive density (elpd_diff):

The top-ranked model in each table has an elpd_diff of 0 and is considered the best supported.

The se_diff column reports the standard error of the difference, quantifying the uncertainty in ranking.

# Models syntax

::: panel-tabset
## Gaussian models (Phenological mismatch)

```{r}
#| label: Gaussian models
#| echo: false
dfg <- read.csv(here("summaries","gaussian.csv"))
DT::datatable(dfg, options = list(pageLength = 5))
```

## Ordinal models (Clutch size and Number of fledglings)

```{r}
#| label: Ordinal models
#| echo: false
dfo <- read.csv(here("summaries","ordinal.csv"))
DT::datatable(dfo, options = list(pageLength = 5))
```
:::

# Phenological Mismatch

The following tables summarize the LOO-CV model comparisons for the Gaussian models of phenological mismatch. Results are shown separately for females and males.

::: panel-tabset
## Females

```{r}
#| label: loo-comparison-pm-females
#| echo: false
#| tbl-cap: "LOO-CV comparison for female phenological mismatch models."

all_comparisonsPMf <- read.csv(here::here("summaries", "all_comparisonsPMf.csv"), header = TRUE) %>%
  select(Model, elpd_diff, se_diff)

DT::datatable(all_comparisonsPMf, options = list(pageLength = 10))
```

## Males

```{r}
#| label: loo-comparison-pm-males
#| echo: false
#| tbl-cap: "LOO-CV comparison for male phenological mismatch models."

all_comparisonsPMm <- read.csv(here::here("summaries", "all_comparisonsPMm.csv"), header = TRUE) %>%
  select(Model, elpd_diff, se_diff)

DT::datatable(all_comparisonsPMm, options = list(pageLength = 10))
```
:::

# Number of Fledglings

The following tables summarize the LOO-CV model comparisons for the ordinal models of fledgling number. Results are shown separately for females and males.

::: panel-tabset
## Females

```{r}
#| label: loo-comparison-fs-females
#| echo: false
#| tbl-cap: "LOO-CV comparison for female fledgling success models."

all_comparisonsFSf <- read.csv(here::here("summaries", "all_comparisonsFSf.csv"), header = TRUE) %>%
  select(Model, elpd_diff, se_diff)

DT::datatable(all_comparisonsFSf, options = list(pageLength = 10))
```

## Males

```{r}
#| label: loo-comparison-fs-males
#| echo: false
#| tbl-cap: "LOO-CV comparison for male fledgling success models."

all_comparisonsFSm <- read.csv(here::here("summaries", "all_comparisonsFSm.csv"), header = TRUE) %>%
  select(Model, elpd_diff, se_diff)

DT::datatable(all_comparisonsFSm, options = list(pageLength = 10))
```
:::

::: {.callout-tip appearance="simple" icon="false"}
```{r}
#| echo: false
#| warning: false
#| message: false
utils::sessionInfo()
```
:::
