---
title: "S6. Full model summaries"
unnumbered: true
---

```{r}
#| label: setup-final-summaries
#| include: false

# Load libraries needed for this chapter
pacman::p_load(here, brms)
```

::: {.callout-note title="Variable definition"}
\-`X5MISMATCH`: Phenological mismatch calculated using the 5% bloom threshold (in days).

\-`CLUTCH_SIZE`: The number of eggs laid in a nest (ordinal).

\-`FLEDS`: The number of chicks that successfully fledged (ordinal).

\-`TIME`: Chronological time, represented as centered years to model temporal trends. Centering is done by subtracting the mean.

\-`ZX5MISMATCH`: The centered phenological mismatch (5% threshold).

\-`AGE`: The age of the focal individual (in years), centered.

\-`AFR`: The age at first reproduction (in years), centered.

\-`LONGEVITY`: The age at the last observation, a proxy for lifespan (in years), centered.

\-`YEAR`: The calendar year of the observation (treated as a factor).

\-`RING`: The unique alphanumeric identifier for each individual (treated as a factor).
:::

This section presents the posterior summaries for the best-supported models from each primary analysis. The models for Phenological mismatch and Number of fledglings were selected based on Leave-One-Out Cross-Validation (LOO-CV)

```{r}
#| label: summary-bloom
#| eval: true
#| size: "footnotesize"


# Load the model
model_bloom <- readRDS(here::here("..","Rdata", "model_delta5.rds"))

# Print the full model summary
summary(model_bloom)

```

::::: panel-tabset
## Females

::: panel-tabset
## Phenological Mismatch

```{r}
#| label: summary-pm-female-final
#| eval: true
#| size: "footnotesize"


# Load the model
model_pm_female_best <- readRDS(here::here("..","Rdata", "female_fit_m8.rds"))

# Print the full model summary
summary(model_pm_female_best)

```

## Number of Fledglings

```{r}
#| label: summary-fs-female-final
#| eval: true
#| size: "footnotesize"


# Load the model
model_fs_female_best <- readRDS(here::here("..","Rdata", "female_fit_Q17.rds"))

# Print the full model summary
print(summary(model_fs_female_best), digits =4)
```
:::

## Males

::: panel-tabset
## Phenological Mismatch

```{r}
#| label: summary-pm-male-final
#| eval: true
#| size: "footnotesize"


# Load the model
model_pm_male_best <- readRDS(here::here("..","Rdata", "male_fit_m8.rds"))

# Print the full model summary
summary(model_pm_male_best)
```

## Number of Fledglings

```{r}
#| label: summary-fs-male-final
#| eval: true
#| size: "footnotesize"


# Load the model
model_fs_male_best <- readRDS(here::here("..","Rdata", "male_fit_Q17.rds"))

# Print the full model summary
print(summary(model_fs_male_best),digits=4)
```
:::
:::::

::: {.callout-tip appearance="simple" icon="false"}
```{r}
#| echo: false
#| warning: false
#| message: false
utils::sessionInfo()
```
:::
