---
title: "Sensitivity analyses: Bloom onset thresholds"
subtitle: "Robustness of bloom phenology inference across threshold definitions"
unnumbered: true
toc: true
toc-depth: 2
---

::: lead
This section evaluates whether inference about bloom phenology is robust to the chlorophyll-a threshold used to define bloom onset. We compare posterior summaries from Gaussian location–scale models fit to bloom onset defined at 5%, 15%, and 25% above the seasonal median chlorophyll-a concentration, and provide the code used to fit these models.
:::

::: {.callout-note appearance="simple"}
## Aim of this sensitivity analysis

These models test whether inference about temporal trends in both the **mean timing** and **variability** of bloom onset remains consistent across alternative bloom-threshold definitions.
:::

```{r}

#| label: setup-threshold-sensitivity
#| include: false

pacman::p_load(
  brms, dplyr, here, future, purrr, furrr, cmdstanr, tibble, stringr
)
```

## Threshold summaries

The tabs below report posterior summaries for the location–scale models fit to bloom onset defined using three alternative chlorophyll-a thresholds. 



::: panel-tabset

## 5% Threshold

```{r}
#| label: summary-5pct
#| eval: true
summary(readRDS(here("..","Rdata","model_delta5.rds")))
```

## 15% Threshold

```{r}
#| label: summary-15pct
#| eval: true
summary(readRDS(here("..","Rdata","model_delta15.rds")))
```

## 25% Threshold

```{r}
#| label: summary-25pct
#| eval: true
summary(readRDS(here("..","Rdata","model_delta25.rds")))
```

:::

## Model fitting

The following code documents the workflow used to transform bloom onset dates and fit the threshold-specific Gaussian location–scale models.

```{r}
#| label: setup-threshold-models
#| eval: false
#| code-fold: true
#| code-summary: "Show computational setup"
#| 
# --- Parallel Processing Setup ---
options(mc.cores = parallel::detectCores(), brms.backend = "cmdstanr")
plan(multisession, workers=4) 
```

### Data transformation

Bloom onset dates were converted to numeric offsets, measured as days since the previous winter solstice (21 December), matching the parameterization used in the main bloom phenology analyses.

```{r}

#| label: threshold-data-transformation
#| eval: false
#| code-fold: true
#| code-summary: "Show data transformation code"
dat <- read.csv(here("..","data","bloom_dates.csv"))

dat <- dat %>%
  mutate(
    # --- Date parsing for detected bloom onsets ---
    bloom_start_thr5_as_date  = as.Date(bloom_start_thr5,  format = "%d/%m/%Y"),
    bloom_start_thr15_as_date = as.Date(bloom_start_thr15, format = "%d/%m/%Y"),
    bloom_start_thr25_as_date = as.Date(bloom_start_thr25, format = "%d/%m/%Y"),

    # --- Reference date: previous winter solstice (Dec 21) ---
    winter_start_date = as.Date(paste0(season_year - 1, "-12-21")),

    # --- Bloom timing response variables (days since winter solstice) ---
    delta5  = as.numeric(difftime(bloom_start_thr5_as_date,  winter_start_date, units = "days")),
    delta15 = as.numeric(difftime(bloom_start_thr15_as_date, winter_start_date, units = "days")),
    delta25 = as.numeric(difftime(bloom_start_thr25_as_date, winter_start_date, units = "days")),

    # --- Year variables with distinct roles ---
    # For summaries / tables / bookkeeping only (not used as the model time trend)
    season_year_f = as.factor(season_year),

    # For location–scale models (mean and sigma trends): centered numeric year
    TIME = as.numeric(scale(season_year, center = TRUE, scale = FALSE)),

    # Optional: chlorophyll values at detected bloom onset (kept as numeric)
    value5  = as.numeric(value_at_bloom_thr5),
    value15 = as.numeric(value_at_bloom_thr15),
    value25 = as.numeric(value_at_bloom_thr25)
  ) %>%
  # Drop original date strings and helper date column after transformation
  select(-starts_with("bloom_start_thr"), -winter_start_date)
```

## Fit Gaussian location–scale models

The same Gaussian location–scale structure was fit to each bloom onset metric (delta5, delta15, delta25), with TIME included in both the location (mean) and scale (residual SD) components.

```{r}
#| label: threshold-location-scale-function
#| eval: false
#| code-fold: true
#| code-summary: "Show location–scale model function"
#| 
run_location_scale_models <- function(response_vars, data) {
  if (!dir.exists(here("Rdata"))) {
    dir.create(here("Rdata"))
  }
  
  # Use purrr::walk to iterate through each response variable and fit a model.
  purrr::walk(response_vars, function(var) {
    
    # Define the model formula. We model both the mean (mu) and the standard
    # deviation (sigma) as a function of TIME.
    model_formula <- bf(
      as.formula(paste(var, "~ 1 + TIME")),
      sigma ~ 1 + TIME
    )
    
    # Define the output file path for the saved model object.
    file_path <- here("Rdata", paste0("model_", var, ".rds"))
    
    # Define the model family.
    model_family <- gaussian()
    
    # Get the default priors for the specified model formula and data.
    model_priors <- get_prior(model_formula, data = data, family = model_family)
    
    # Use tryCatch to gracefully handle any potential errors during model fitting.
    fit <- tryCatch({
      brm(
        formula = model_formula,
        data = data,
        prior = model_priors,
        family = model_family,
        iter = 6000,
        warmup = 2000,
        chains = 4,
        seed = 123,
        init = 0,
        control = list(adapt_delta = 0.99, max_treedepth = 15),
        file = file_path,
        file_refit = "on_change"
      )
    }, error = function(e) {
      # If an error occurs, print a message and return NULL.
      cat("Error fitting model for:", var, "\n")
      cat("Error message:", e$message, "\n")
      return(NULL)
    })
    
    # Only print the success message if the model fitting was successful.
    if (!is.null(fit)) {
      cat("Saved model for:", var, "to", file_path, "\n")
    }
  })
}
```

### Model execution

```{r}
#| label: run-models
#| eval: false
# --- Run Analysis for Delta Variables ---
# Define the list of delta response variables.
delta_vars <- c("delta5", "delta15", "delta25")
# Run the models for the delta variables.
run_location_scale_models(response_vars = delta_vars, data = dat)
```

## Reproducibility

::: {.callout-tip appearance="simple" icon="false"}
## Session information

```{r}
#| echo: false
#| warning: false
#| message: false
utils::sessionInfo()
```
:::
